Applying gamification to developer practices

### 2013-09-10 paper_id1.tex
anil's update

\documentclass{sig-alternate}
% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

@@

Maehr proposes an affirmative theory that individuals achieve as a member of a social group choosing the behavior that will meet the expectations and values of the group that is significant to them. \cite{wbsnipes:MaehrCulture}  So the presence of a social group where the person is motivated to belong and excel within is important for creating achievement motivation.  Maehr states that since achievement behaviors can be triggered by circumstance, that they are present in most people regardless of background.

Beecham and colleagues provide a thorough analysis of existing literature on studies of motivation in Software Engineering\cite{wbsnipes:Beecham2008Motivation}. It has some valuable lists of factors that motivate software engineers with the most common being "the work" motivates us. The list of de-motivators is also useful with common job satisfaction items like stress and inequity in recognition, plus poor quality software (low accomplishment) and lack of influence in decision making. The paper also lists characteristics of people in the professions including need for independence (autonomy), desire to learn new skills/tackle challenges.

@@

Using a question from the Saachi and Saachi survey\cite{wbsnipes:SaatchiGameification} as directly as possible, we asked developers at ABB "How interested they would be in working for a company that incorporated some aspects of games into software engineering tools as a way to increase productivity in the workplace"?  We segmented results by country in Figure \ref{fig:gamification} and found that gamification of software engineering is more interesting to developers in India, Poland, and Finland than in other major countries.  The responses indicating that developers were at least somewhat interested is 75-88\% In India, Poland, and Finland.  For the US, Sweden, and Switzerland the responses indicate 60-65\% are at least somewhat interested.  The overall ABB response is comparable to the Saachi and Saaschi survey\cite{wbsnipes:SaatchiGameification} upon which this question was based that found 75\% of 18-45 year olds were at least somewhat interested.

@@

Based on the survey results we made the following conclusions:
•	95\% of developer responding would try tools and practices suggested by an automated recommendation system particularly if they were provided in context with the current workflow pattern the developer follows.
•	Responses indicate a preference for team competition over individual awards (badges) and awards for quality practices ranked highest.  So, the tool will incorporate team competition ideas and recognize quality practices.
•	Over 90\% of respondents approved of sharing data with their team members

@@

Utilizing the unobtrusive continuous monitoring capabilities of Blaze, we designed the pilot as a longtiudinal study in three waves.  Longtiudinal studies allow us to design an experiment and determine the effects over two or more sample periods using standard MANOVA (Multiple ANalysis of VAriance) models. \cite{longitudinalbook}  We employed the three wave model defining the first wave as a baseline capturing the current practices of the pilot group.  The second wave employed the intervention that included an informative web site on structured navigation commands in Visual Studio, information about the Sando search tool, and information about how points would be awarded during the pilot.  The third wave provided the participants with continuous feedback in the from of points accumulated and an indicator arrow \ref{fig:blazegui} that points up when their navigation ratio is higher than their historical average.

Developing the study with the available tooling means that we follow steps as follows.
1. Define the practice we wish to measure compliance with
2. Develop a mapping of the practice to actions that are captured by the blaze tool from the IDE
3. Weight the actions according to their importance to the practice and assign points to the actions in the configuration file
4. Develop the intervention method
5. Determine the phases and length of the phase in the study according to baseline, intervention, and continuous feedback
6. Select participants to deploy
7. Define a pre-study questionnaire if one is necessary to collect associated information
8. Define a post-study questionnaire if necessary to collect post study information
9. Communicate with the participants about the study
10. Execute the study and report results

@@

We conducted this study with an intact team of 6 developers working in an R\&D facility in India on a large industrial software system. We approached a team which develops and maintains an industrial software product. Initially, we met the whole team along with the management staff to explain the research project and the Blaze tool we developed. We  explained the aim of the study, what data is collected, and how the data will be processed and managed.

It was important to gain the confidence of the participants that their data were kept confidential so their activities would be close as possible to real-world. if the participants felt like they are being tracked and the data is going to use for some study that would identify them to their peers or managers, then they may deviate from the natural development style.

@@

Competitive dynamics may be more effective when teams from different areas can compete as a team against other teams..  This tried and true method both builds esprit de corps and enhances achievement towards the intended outcome.  This study design is potential future work.

@@

The following hypotheses interest us in the current study.  
1. Developers using blaze use more structured navigation when they receive instant feedback about their practices
2. Developers find the competitive aspect of blaze encourages them to pay more attention to navigation practices

@@

Is this a spectacular fail from a gamification effectiveness perspective?   How could this be when pre-study results indicated very positive attitudes towards the approach?  Our post pilot survey attempts to shed some light on the reasons.  One question in the post survey is how knowledgeable the developer was about the code they worked in with 3 possibilities,  they were authors of the code, they regularly maintained the code, they were unfamiliar with the code.  In this team, x\% of developers cited that they ...the code.  Another factor to consider is experience level which we also asked in the post-pilot survey.  More experienced developers may be more convinced of their effective habits and perhaps react almost with muscle memory to use structured or unstructured navigation to locate code of interest.  Even developers whose navigation ratio  was lower than average, used structured navigation occasionally as much as their more frequently using colleagues.  The variation for each individual can be seen in Figure \ref{} .

>>>>

\documentclass{sig-alternate}
% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage{enumerate}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

@@

Maehr \cite{wbsnipes:MaehrCulture} proposes an affirmative theory that individuals achieve as a member of a social group choosing the behavior that will meet the expectations and values of the group that is significant to them.   So the presence of a social group where the person is motivated to belong and excel within is important for creating achievement motivation.  Maehr states that since achievement behaviors can be triggered by circumstance, they are present in most people regardless of background.

Beecham and colleagues provide a thorough analysis of existing literature on studies of motivation in Software Engineering\cite{wbsnipes:Beecham2008Motivation}. It has some valuable lists of factors that motivate software engineers with the most common being ``the work'' motivates us. The list of de-motivators is also useful with common job satisfaction items like stress and inequity in recognition, plus poor quality software (low accomplishment) and lack of influence in decision making. The paper also lists characteristics of people in the professions including need for independence (autonomy), desire to learn new skills/tackle challenges.

@@

Using a question from the Saachi and Saachi survey\cite{wbsnipes:SaatchiGameification} as directly as possible, we asked developers at ABB "How interested they would be in working for a company that incorporated some aspects of games into software engineering tools as a way to increase productivity in the workplace"?  We segmented results by country in Figure \ref{fig:gamification} and found that gamification of software engineering is more interesting to developers in India, Poland, and Finland than in other major countries.  The responses indicating that developers were at least somewhat interested is 75-88\% in India, Poland, and Finland.  For the US, Sweden, and Switzerland the responses indicate 60-65\% are at least somewhat interested.  The overall ABB response is comparable to the Saachi and Saaschi survey\cite{wbsnipes:SaatchiGameification} upon which this question was based that found 75\% of 18-45 year olds were at least somewhat interested.

@@

Based on the survey results we made the following conclusions:

\begin{itemize}
 \item 95\% of developer responding would try tools and practices suggested by an automated recommendation system particularly if they were provided in context with the current workflow pattern the developer follows.
\item Responses indicate a preference for team competition over individual awards (badges) and awards for quality practices ranked highest.  So, the tool will incorporate team competition ideas and recognize quality practices.
\item Over 90\% of respondents approved of sharing data with their team members
\end{itemize}

@@

Utilizing the unobtrusive continuous monitoring capabilities of Blaze, we designed the pilot as a longitudinal study in three waves.  Longtiudinal studies allow us to design an experiment and determine the effects over two or more sample periods using standard MANOVA (Multiple ANalysis of VAriance) models. \cite{longitudinalbook}  We employed the three wave model defining the first wave as a baseline capturing the current practices of the pilot group.  The second wave employed the intervention that included an informative web site on structured navigation commands in Visual Studio, information about the Sando search tool, and information about how points would be awarded during the pilot.  The third wave provided the participants with continuous feedback in the from of points accumulated and an indicator arrow \ref{fig:blazegui} that points up when their navigation ratio is higher than their historical average.

Developing the study with the available tooling means that we follow steps as follows.

\begin{enumerate}
\item Define the practice we wish to measure compliance with
\item Develop a mapping of the practice to actions that are captured by the blaze tool from the IDE
\item Weight the actions according to their importance to the practice and assign points to the actions in the configuration file
\item Develop the intervention method
\item Determine the phases and length of the phase in the study according to baseline, intervention, and continuous feedback
\item Select participants to deploy
\item Define a pre-study questionnaire if one is necessary to collect associated information
\item Define a post-study questionnaire if necessary to collect post study information
\item Communicate with the participants about the study
\item Execute the study and report results
\end{enumerate}

@@

As mentioned in fig. \ref{fig:gamification}, India is one country where as per our survey, the acceptance to gamification is the highest. Hence, we conducted this study with an intact team of 6 developers working in an R\&D facility in India on a large industrial software system. We approached a team which develops and maintains an industrial software product. Initially, we met the whole team along with the management staff to explain the research project and the Blaze tool we developed. We  explained the aim of the study, what data is collected, and how the data will be processed and managed.

It was important to gain the confidence of the participants that their data were kept confidential so their activities would be close as possible to real-world. If the participants felt like they are being tracked and the data is going to use for some study that would identify them to their peers or managers, then they may deviate from the natural development style.

@@

Competitive dynamics may be more effective when teams from different areas can compete as a team against other teams.  This tried and true method both builds esprit de corps and enhances achievement towards the intended outcome.  This study design is potential future work.

@@

The following hypotheses interest us in the current study.

\begin{enumerate}
\item Developers using blaze use more structured navigation when they receive instant feedback about their practices
\item Developers find the competitive aspect of blaze encourages them to pay more attention to navigation practices
\end{enumerate}

@@

Is this a spectacular fail from a gamification effectiveness perspective?   How could this be when pre-study results indicated very positive attitudes towards the approach?  Our post pilot survey attempts to shed some light on the reasons.  One question in the post survey is how knowledgeable the developer was about the code they worked in with 3 possibilities,  they were authors of the code, they regularly maintained the code, they were unfamiliar with the code.  In this team,67\% of developers cited that they work with the code they wrote and the balance 33\% works with the code they recently learned.  Another factor to consider is experience level which we also asked in the post-pilot survey.  More experienced developers may be more convinced of their effective habits and perhaps react almost with muscle memory to use structured or unstructured navigation to locate code of interest.  Even developers whose navigation ratio  was lower than average, used structured navigation occasionally as much as their more frequently using colleagues.  The variation for each individual can be seen in Figure \ref{} .
